---
title: "ESM_201_A_2"
output:
  pdf_document: default
  html_notebook: default
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

1a)

```{r}

# attach packages
library(tidyverse)
library(janitor)

grains <- read_csv("grain_2020.csv") %>% 
  clean_names()
  
##

ggplot(data = grains, aes(x = year, y = yield)) + 
         geom_line(aes(color = commodity)) +
  facet_wrap(~ commodity, scales = "free") +
  labs(x = "Years", y = "Yield (Bushels per hectare)", title = "Crop Yields by Time", caption = "Figure 1: Graph of crop yields \n (Bushels per hectare) of Barley, Corn, Oats, and Sorghum by time") +
  theme(plot.caption = element_text(hjust = 0))

  


```


#1b)
According to the graphs, it looks as if Barley, Oats, and Corn  are Linear, Lower Plateau(LLP). However, Sorghum looks like it could be Linear, upper plateau(LUP) or maybe even Linear piecewise (PW) statistic model. 

```{r}
grain_extra <- read_csv("grain_2020.csv") %>% 
  clean_names() %>% 
  rename(crop = commodity) %>% 
  filter(crop == "Corn") 




ggplot(data = grain_extra, aes(x = year, y = yield)) + 
         geom_line() +
  labs(x = "Years", y = "Yield (Bushels per hectare)", title = "Corn Yields by Time", caption = "Figure 1b: Graph of Corn yields \n (Bushels per hectare) by time") +
  theme(plot.caption = element_text(hjust = 0))


```

#2a)

```{r}

fert <- read_csv("fertilizer_2020.csv") %>% 
  clean_names()
  
##

ggplot(data = fert, aes(x = year, y = kg_h)) + 
         geom_line(aes(color = fertilizer)) +
  facet_wrap(~ crop) +
  labs(x = "Years", y = "Fertilizer Use (kilogram per hectare)", title = "Fertilizer Use by Time", caption = "Figure 2: Graph of fertilizer Use (kilogram per hectare) \n of Wheat, Soybeans, and Corn by time") +
  theme(plot.caption = element_text(hjust = 0)) +
   scale_x_continuous(limits = c(1967, 2006))

```



2b)
Observations:
Corn and Wheat have more Nitrogen fertilizer than the other fertilizers. Additionally, Corn has significantly more nitrogen use than any of the other two crops. Soybeans has more potassium than nitrogen and phosphate.

For Corn, the limiting nutrient is Nitrogen because it is needed the most for the yields to grow. For Soybeans, the limiting nutrient is Potassium. For Wheat, the limiting nutrient is Nitrogen as well.

3a)


```{r}


# rename commodity to crop 
grain_c <- read_csv("grain_2020.csv") %>% 
  clean_names() %>% 
  rename(crop = commodity) %>% 
  filter(crop == "Corn") %>% 
  filter(year %in% c("1964":"2016")) %>% 
  group_by(year)

fert_c <- read_csv("fertilizer_2020.csv") %>% 
  clean_names() %>% 
  filter(crop == "Corn")
  

#join tables
fert_grain <- full_join(grain_c,fert_c, by = "year") 
  

ggplot(data = fert_grain, aes(x = yield, y = kg_h)) + 
         geom_point(aes(color = fertilizer)) +
  geom_smooth(aes(color = fertilizer)) +
  labs(x = "Yield", y = "Fertilizer Use (kilogram per hectare)", title = "Fertilizer Use by Yield for Corn", caption = "Figure 3: Graph of fertilizer Use (kilogram per hectare) \n of Corn by yield") +
 theme(plot.caption = element_text(hjust = 0)) 
  



```

3b) Nitrogen fertilizer use is needed in high amounts of 160+ kg per hectare to get to more than 400 Bushels per hectare of corn yield. However, you do not need that much Phosphate and Potassium to reach the same amount of yields. As you increase yield, you do not need additional Phosphate and Potassium. But, you do need a lot more Nitrogen as you want to increase yields.

The relationship is not linear. Although it does yield does increase in the beginning as you add fertilizer, it tapers out. It no longer provides high yields when you add more fertilizer.
